### 2014-09-11 paper/paper.Rmd
fixed code error, and axis labels

```{r age, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
student2012.sub <- student2012[, c(1, 6, 12, 14, 501, 541, 546, 634)]
student2012.sub.age <- summarise(group_by(student2012.sub, name), 
                  math4 = mean(PV1MATH[ST06Q01 == 4], na.rm = T), 
                  math5 = mean(PV1MATH[ST06Q01 == 5], na.rm = T), 
                  math6 = mean(PV1MATH[ST06Q01 == 6], na.rm = T), 
                  math7 = mean(PV1MATH[ST06Q01 == 7], na.rm = T), 
                  math8 = mean(PV1MATH[ST06Q01 >  7], na.rm = T))
student2012.sub.age.m <- melt(student2012.sub.age)
student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
student2012.sub.agemedian <- summarise(group_by(student2012.sub, name), medianstart = median(ST06Q01, na.rm=T))
orderedbymedianage <- student2012.sub.agemedian$name[order(student2012.sub.agemedian$medianstart)]
student2012.sub.age.m.median <- left_join(student2012.sub.age.m, student2012.sub.agemedian)
student2012.sub.age.m.median$medianstart <- factor(student2012.sub.age.m.median$medianstart)
student2012.sub.age.m.median$name <- factor(student2012.sub.age.m.median$name, levels=orderedbymedianage)
qplot(variable, value, data=student2012.sub.age.m.median, xlab="Age at start", ylab="Math Scores", 
      geom=c("point","smooth"), se=F, group=name, colour=medianstart) + 
 facet_wrap(~name, ncol=8) + theme_bw() + theme(legend.position="none")
# ggsave("school-start.pdf", width=14, height=14)
#student2012.sub.age <- summarise(group_by(student2012.sub, name), 
#                  read4 = mean(PV1READ[ST06Q01 == 4], na.rm = T), 
#                  read5 = mean(PV1READ[ST06Q01 == 5], na.rm = T), 
#                  read6 = mean(PV1READ[ST06Q01 == 6], na.rm = T), 
#                  read7 = mean(PV1READ[ST06Q01 == 7], na.rm = T), 
#                  read8 = mean(PV1READ[ST06Q01 >  7], na.rm = T))
#student2012.sub.age.m <- melt(student2012.sub.age)
#student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
#student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
#qplot(variable, value, data=student2012.sub.age.m, xlab="Age at start", ylab="Reading Score", 
#      geom=c("point","smooth"), se=F, group=name) + 
# facet_wrap(~name, ncol=8) + theme_bw() 
#student2012.sub.age <- summarise(group_by(student2012.sub, name), 
#                  scie4 = mean(PV1SCIE[ST06Q01 == 4], na.rm = T), 
#                  scie5 = mean(PV1SCIE[ST06Q01 == 5], na.rm = T), 
#                  scie6 = mean(PV1SCIE[ST06Q01 == 6], na.rm = T), 
#                  scie7 = mean(PV1SCIE[ST06Q01 == 7], na.rm = T), 
#                  scie8 = mean(PV1SCIE[ST06Q01 >  7], na.rm = T))
#student2012.sub.age.m <- melt(student2012.sub.age)
#student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
#student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
#qplot(variable, value, data=student2012.sub.age.m, xlab="Age at start", ylab="Science Score", 
#      geom=c("point","smooth"), se=F, group=name) + 
# facet_wrap(~name, ncol=8) + theme_bw() 
```

```{r agemaps, fig.width=12, fig.height=7, echo = FALSE, message = FALSE, warning = FALSE}
student2012.sub.agemedian$medianstart <- factor(student2012.sub.agemedian$medianstart)
student2012.sub.map <- left_join(student2012.sub.agemedian, world.polys)
ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey90")) + 
  geom_polygon(data=student2012.sub.map, aes(x=X1, y=X2, order=order, group=group, fill=medianstart)) +
  scale_fill_discrete("Median Start Age") +
  new_theme_empty + theme(legend.position="bottom")
#ggsave("agestartmap.pdf", width=12, height=7)
```

@@

```{r shortagesbars, fig.width=10, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
df$SC14Q02 <- as.character(df$SC14Q02)
df$SC14Q07 <- as.character(df$SC14Q07)
df$SC14Q10 <- as.character(df$SC14Q10)
df$SC14Q02[df$SC14Q02=="A lot"] <- "To some extent"
df$SC14Q07[df$SC14Q07=="A lot"] <- "To some extent"
df$SC14Q10[df$SC14Q10=="A lot"] <- "To some extent"
df$SC14Q02 <- factor(df$SC14Q02, labels=c(0,1,2))
df$SC14Q07 <- factor(df$SC14Q07, labels=c(0,1,2))
df$SC14Q10 <- factor(df$SC14Q10, labels=c(0,1,2))
df.mshort <- df[!is.na(df$SC14Q02),] %>% group_by(name, SC14Q02) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q02, mmath2, data = df.mshort, facets = ~name, 
#      xlab="Math teacher shortage (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df.lshort <- df[!is.na(df$SC14Q10),] %>% group_by(name, SC14Q10) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q10, mmath2, data = df.lshort, facets = ~name, 
#      xlab="Library shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df2 <- summarise(group_by(df[!is.na(df$SC14Q02),], name, SC14Q02), teachers = mean(mmath, na.rm=T))
colnames(df2)[2] <- "levels"
df3 <- summarise(group_by(df[!is.na(df$SC14Q07),], name, SC14Q07), computers = mean(mmath, na.rm=T))
colnames(df3)[2] <- "levels"
df4 <- summarise(group_by(df[!is.na(df$SC14Q10),], name, SC14Q10), libmaterials = mean(mmath, na.rm=T))
colnames(df4)[2] <- "levels"
df.all <- merge(df2, df3)
df.all <- merge(df.all, df4)
df.m <- melt(df.all, id=c("name", "levels"))
colnames(df.m)[4] <- "math"
qplot(levels, math, data = df.m, facets = ~name, 
      xlab="Shortages (0=None, 1=Very little, 2=To some extent, or A lot)", 
      ylab="Math Score", geom=c("point","line"), group=variable, colour=variable) +
      scale_colour_manual("Type", values=c("teachers"="#1B9E77", "computers"="#D95F02", "libmaterials"="#7570B3")) +
      theme(legend.position="bottom") 
```

@@

```{r studying, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
student.luke <- student2012[,c("name","PV1MATH","PV1READ","PV1SCIE","ST57Q01")]
#study.luke <- melt(student.luke,id=c("name","ST57Q01"))
studyplot <- ggplot(data = student.luke, aes(ST57Q01, PV1MATH), xlab="Time spent", ylab="Math scores") + facet_wrap(~name) + stat_smooth(se=F)
studyplot
```

@@

```{r par_athome, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
phome <- qplot(x = parents, y = PV1MATH, geom = "boxplot", outlier.size = 1, data = athome) + ylab("Math") + facet_wrap(~name, scales="free_y") + scale_x_discrete("Family Structure", labels=c("Mother"="M", "Father"="F", "Both"="B", "Neither"="N")) 
phome
```

@@

a8 <- rbind(cbind(a6,Parent="Mother"),cbind(a7,Parent="Father"))
#a2 <- melt(a1)
ind <- which(a8$Job.Status=="Full-Time"& a8$Parent=="Mother")
a8$name <- factor(a8$name,levels(a8$name)[order(a8$Math[ind])])
p <- qplot(name, Math, col=Job.Status, data = a8, xlab="") + coord_flip() + facet_wrap(~Parent) + theme(legend.position="bottom")
p
```

@@

#df$group = paste0(df$name, df$variable)
#qplot(x = numposs, y = value, data = df, facets = ~OECD, geom = "line", color = variable, group = group) + ylab("test scores") + xlab("number of possessions") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
qplot(x = numposs, y = value, data = df, facets = ~name, geom = "line") + ylab("Math Scores") + xlab("Number of Possessions") + theme_bw()
```

@@

#qplot(x = SC22Q01, y = value, color = variable, facets = ~name, data = truancy.df, group = group, geom = "line") + xlab("Reported Truancy") + ylab("Test Scores") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
qplot(x = SC22Q01, y = value, facets = ~name, data = truancy.df, group=name, geom="smooth", se=F) + 
  scale_x_discrete("Reported Truancy, 0=None, 1=Little, 2=Some, 3=Lot", labels=c("Not"="0", "Little"="1", "Some"="2", "Lot"="3")) + 
  ylab("Math Scores") + theme_bw()
```

>>>>

```{r age, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
student2012.sub <- student2012[, c(1, 6, 12, 14, 501, 541, 546, 634)]
student2012.sub.age <- summarise(group_by(student2012.sub, name), 
                  math4 = mean(PV1MATH[ST06Q01 == 4], na.rm = T), 
                  math5 = mean(PV1MATH[ST06Q01 == 5], na.rm = T), 
                  math6 = mean(PV1MATH[ST06Q01 == 6], na.rm = T), 
                  math7 = mean(PV1MATH[ST06Q01 == 7], na.rm = T), 
                  math8 = mean(PV1MATH[ST06Q01 >  7], na.rm = T))
student2012.sub.age.m <- melt(student2012.sub.age)
student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
student2012.sub.agemedian <- summarise(group_by(student2012.sub, name), medianstart = median(ST06Q01, na.rm=T))
orderedbymedianage <- student2012.sub.agemedian$name[order(student2012.sub.agemedian$medianstart)]
student2012.sub.age.m.median <- left_join(student2012.sub.age.m, student2012.sub.agemedian)
student2012.sub.age.m.median$medianstart <- factor(student2012.sub.age.m.median$medianstart)
student2012.sub.age.m.median$name <- factor(student2012.sub.age.m.median$name, levels=orderedbymedianage)
qplot(variable, value, data=student2012.sub.age.m.median, xlab="Age at start", ylab="Math score", 
      geom=c("point","smooth"), se=F, group=name, colour=medianstart) + 
 facet_wrap(~name, ncol=8) + theme_bw() + theme(legend.position="none")
# ggsave("school-start.pdf", width=14, height=14)
#student2012.sub.age <- summarise(group_by(student2012.sub, name), 
#                  read4 = mean(PV1READ[ST06Q01 == 4], na.rm = T), 
#                  read5 = mean(PV1READ[ST06Q01 == 5], na.rm = T), 
#                  read6 = mean(PV1READ[ST06Q01 == 6], na.rm = T), 
#                  read7 = mean(PV1READ[ST06Q01 == 7], na.rm = T), 
#                  read8 = mean(PV1READ[ST06Q01 >  7], na.rm = T))
#student2012.sub.age.m <- melt(student2012.sub.age)
#student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
#student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
#qplot(variable, value, data=student2012.sub.age.m, xlab="Age at start", ylab="Reading Score", 
#      geom=c("point","smooth"), se=F, group=name) + 
# facet_wrap(~name, ncol=8) + theme_bw() 
#student2012.sub.age <- summarise(group_by(student2012.sub, name), 
#                  scie4 = mean(PV1SCIE[ST06Q01 == 4], na.rm = T), 
#                  scie5 = mean(PV1SCIE[ST06Q01 == 5], na.rm = T), 
#                  scie6 = mean(PV1SCIE[ST06Q01 == 6], na.rm = T), 
#                  scie7 = mean(PV1SCIE[ST06Q01 == 7], na.rm = T), 
#                  scie8 = mean(PV1SCIE[ST06Q01 >  7], na.rm = T))
#student2012.sub.age.m <- melt(student2012.sub.age)
#student2012.sub.age.m$variable <- substr(student2012.sub.age.m$variable, 5, 5)
#student2012.sub.age.m$variable <- as.numeric(student2012.sub.age.m$variable)
#qplot(variable, value, data=student2012.sub.age.m, xlab="Age at start", ylab="Science Score", 
#      geom=c("point","smooth"), se=F, group=name) + 
# facet_wrap(~name, ncol=8) + theme_bw() 
```

```{r agemaps, fig.width=12, fig.height=7, echo = FALSE, message = FALSE, warning = FALSE}
student2012.sub.agemedian$medianstart <- factor(student2012.sub.agemedian$medianstart)
student2012.sub.map <- left_join(student2012.sub.agemedian, world.polys)
ggplot(data=world.polys) + geom_path(aes(x=X1, y=X2, order=order, group=group), colour=I("grey90")) + 
  geom_polygon(data=student2012.sub.map, aes(x=X1, y=X2, order=order, group=group, fill=medianstart)) +
  scale_fill_discrete("Median start age") +
  new_theme_empty + theme(legend.position="bottom")
#ggsave("agestartmap.pdf", width=12, height=7)
```

@@

```{r shortagesbars, fig.width=10, fig.height=12, warning=FALSE, message=FALSE, echo=FALSE}
df$SC14Q02 <- as.character(df$SC14Q02)
df$SC14Q07 <- as.character(df$SC14Q07)
df$SC14Q10 <- as.character(df$SC14Q10)
df$SC14Q02[df$SC14Q02==3] <- 2
df$SC14Q07[df$SC14Q07==3] <- 2
df$SC14Q10[df$SC14Q10==3] <- 2
df$SC14Q02 <- factor(df$SC14Q02, labels=c(0,1,2))
df$SC14Q07 <- factor(df$SC14Q07, labels=c(0,1,2))
df$SC14Q10 <- factor(df$SC14Q10, labels=c(0,1,2))
df.mshort <- df[!is.na(df$SC14Q02),] %>% group_by(name, SC14Q02) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q02, mmath2, data = df.mshort, facets = ~name, 
#      xlab="Math teacher shortage (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df.lshort <- df[!is.na(df$SC14Q10),] %>% group_by(name, SC14Q10) %>% summarise(mmath2 = mean(mmath, na.rm=T))
#qplot(SC14Q10, mmath2, data = df.lshort, facets = ~name, 
#      xlab="Library shortages (0=None, 1=Very little, 2=To some extent, 3=A lot)", 
#      ylab="Math Score", geom=c("point","line"), group=name)
df2 <- summarise(group_by(df[!is.na(df$SC14Q02),], name, SC14Q02), teachers = mean(mmath, na.rm=T))
colnames(df2)[2] <- "levels"
df3 <- summarise(group_by(df[!is.na(df$SC14Q07),], name, SC14Q07), computers = mean(mmath, na.rm=T))
colnames(df3)[2] <- "levels"
df4 <- summarise(group_by(df[!is.na(df$SC14Q10),], name, SC14Q10), libmaterials = mean(mmath, na.rm=T))
colnames(df4)[2] <- "levels"
df.all <- merge(df2, df3)
df.all <- merge(df.all, df4)
df.m <- melt(df.all, id=c("name", "levels"))
colnames(df.m)[4] <- "math"
qplot(levels, math, data = df.m, facets = ~name, 
      xlab="Shortages (0=None, 1=Very little, 2=To some extent, or a lot)", 
      ylab="Math score", geom=c("point","line"), group=variable, colour=variable) +
      scale_colour_manual("Type", values=c("teachers"="#1B9E77", "computers"="#D95F02", "libmaterials"="#7570B3")) +
      theme(legend.position="bottom") 
```

@@

```{r studying, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
student.luke <- student2012[,c("name","PV1MATH","PV1READ","PV1SCIE","ST57Q01")]
#study.luke <- melt(student.luke,id=c("name","ST57Q01"))
studyplot <- ggplot(data = student.luke, aes(ST57Q01, PV1MATH)) + xlab("Hours spent out of school studying per week") + ylab("Math score") + facet_wrap(~name) + stat_smooth(se=F)
studyplot
```

@@

```{r par_athome, fig.width=10, fig.height=12, echo = FALSE, message = FALSE, warning = FALSE}
phome <- qplot(x = parents, y = PV1MATH, geom = "boxplot", outlier.size = 1, data = athome) + ylab("Math score") + facet_wrap(~name, scales="free_y") + scale_x_discrete("Family Structure", labels=c("Mother"="M", "Father"="F", "Both"="B", "Neither"="N")) 
phome
```

@@

a8 <- rbind(cbind(a6,Parent="Mother"),cbind(a7,Parent="Father"))
#a2 <- melt(a1)
ind <- which(a8$Job.Status=="Full-Time"& a8$Parent=="Mother")
a8$name <- factor(a8$name,levels(a8$name)[order(a8$Math[ind])])
p <- qplot(name, Math, col=Job.Status, data = a8, xlab="Math score") + coord_flip() + facet_wrap(~Parent) + theme(legend.position="bottom")
p
```

@@

#df$group = paste0(df$name, df$variable)
#qplot(x = numposs, y = value, data = df, facets = ~OECD, geom = "line", color = variable, group = group) + ylab("test scores") + xlab("number of possessions") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
qplot(x = numposs, y = value, data = df, facets = ~name, geom = "line") + ylab("Math score") + xlab("Number of possessions") + theme_bw()
```

@@

#qplot(x = SC22Q01, y = value, color = variable, facets = ~name, data = truancy.df, group = group, geom = "line") + xlab("Reported Truancy") + ylab("Test Scores") + scale_colour_discrete(name = "Subject", labels=c("Math", "Science", "Reading"))
qplot(x = SC22Q01, y = value, facets = ~name, data = truancy.df, group=name, geom="smooth", se=F) + 
  scale_x_discrete("Reported truancy, 0=None, 1=Little, 2=Some, 3=Lot", labels=c("Not"="0", "Little"="1", "Some"="2", "Lot"="3")) + 
  ylab("Math score") + theme_bw()
```
